Spin-orbit-coupling-induced squeezing in the optical lattice clock

### 2018-07-19 soc_squeezing/soc_squeezing_notes.tex
added example figure for squeezing in 2D

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.59$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

For small system sizes, we can verify the validity of the effective
one- and two-axis twisting Hamiltonians via direct simulations of a
Fermi-Hubbard (FH) model, using the spin-orbit coupling Hamiltonian
$H_{\t{SOC}}$ and interaction Hamiltonian $H_{\t{int}}$ respectively
given in \eqref{eq:H_SOC_start} and \eqref{eq:H_int_full}.  Figure
\ref{fig:squeezing_comparison_FH} shows a comparison of the squeezing
parameter $\xi^2$ after evolution of the initial state $\ket{Y}$ for
$N=8$ particles in a 1-D lattice with $L=8$ sites (and all remaining
parameters as given in Table \ref{tab:parameters}) after evolution
under
\begin{enumerate*}[label=(\roman*)]
\item the one-axis twisting Hamiltonian $H_{\t{OAT}}^z$,
\item the two-axis twisting Hamiltonian $H_{\t{OAT}}^{z,x}$,
\item free evolution of the Fermi-Hubbard model under
  $H_{\t{SOC}}+H_{\t{int}}$, and
\item driven evolution of the Fermi-Hubbard model using the
  continuous-drive protocol in Section \ref{sec:continuous_drive} with
  drive frequency $\omega=\abs{\eta U\chi}^{1/2}$ and modulation index
  $\beta=0.906$.
\end{enumerate*}
These simulation results show essentially perfect agreement between
the Fermi-Hubbard model and the one- and two-axis twisting
Hamiltonians.

\begin{figure}
  \centering \includegraphics{squeezing_comparison_FH.pdf}
  \caption{Benchmarking squeezing under the effective one- and
    two-axis Hamiltonians $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$
    against direct simulations of the Fermi-Hubbard model with
    $N=L=8$.}
  \label{fig:squeezing_comparison_FH}
\end{figure}

% \note{[MP] Future ideas:} Make use of a cavity to enhance the
% one-axis twisting strength \cite{hu2017vacuum}?

>>>>

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
    SOC strength & $\phi$ & $\pi/50$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.59$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

For small system sizes, we can verify the validity of the effective
one- and two-axis twisting Hamiltonians via direct simulations of a
single-band Fermi-Hubbard (FH) model, using the spin-orbit coupling
Hamiltonian $H_{\t{SOC}}$ and interaction Hamiltonian $H_{\t{int}}$
respectively given in \eqref{eq:H_SOC_start} and
\eqref{eq:H_int_full}.  Figure \ref{fig:squeezing_comparison_FH} shows
a comparison of the squeezing parameter $\xi^2$ after evolution of the
initial state $\ket{Y}$ for $N=8$ particles in a 1-D lattice with
$L=8$ sites (and all remaining parameters as given in Table
\ref{tab:parameters}) after evolution under
\begin{enumerate*}[label=(\roman*)]
\item the one-axis twisting Hamiltonian $H_{\t{OAT}}^z$,
\item the two-axis twisting Hamiltonian $H_{\t{OAT}}^{z,x}$,
\item free evolution of the Fermi-Hubbard model under
  $H_{\t{SOC}}+H_{\t{int}}$, and
\item driven evolution of the Fermi-Hubbard model using the
  continuous-drive protocol in Section \ref{sec:continuous_drive} with
  drive frequency $\omega=\abs{\eta U\chi}^{1/2}$ and modulation index
  $\beta=0.906$.
\end{enumerate*}
These simulation results show essentially perfect agreement between
the Fermi-Hubbard model and the one- and two-axis twisting
Hamiltonians for the given parameters.

\begin{figure}
  \centering
  \includegraphics{squeezing_comparison_FH.pdf}
  \caption{Benchmarking squeezing under the one- and two-axis
    Hamiltonians $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ against
    direct simulations of the Fermi-Hubbard model with $N=L=8$ and the
    remaining lattice/SOC parameters as given in Table
    \ref{tab:parameters}.}
  \label{fig:squeezing_comparison_FH}
\end{figure}

\section{Last thoughts}

It may be possible to further enhance the twisting strength $\chi$ by
placing the atoms in an optical cavity, as in
ref.~\cite{hu2017vacuum}.  Both squeezing times and squeezing
strengths might also be further enhanced by using a 2-D lattice,
although more benchmarking is necessary to verify the validity of the
collective spin model in 2-D.  As shown in Figure
\ref{fig:squeezing_comparison_2D},

\begin{figure}
  \centering
  \includegraphics{squeezing_comparison_2D.pdf}
  \caption{Squeezing via one- and two-axis twisting in a 2-D lattice
    with $100\times100$ sites at half filling, with the remaining
    lattice/SOC parameters given in Table \ref{tab:parameters}.}
  \label{fig:squeezing_comparison_2D}
\end{figure}
